Probabilistic Inference of Basic Stellar Parameters:\\ Application to Flickering Stars \altaffilmark{\titledag}

### 2015-06-05 version1.0/HBM.tex
making some corrections to the text

We model the stochastic relationship between flicker, log($g$) and stellar
density, accounting for the fact that there exists some intrinsic scatter in
the dependent variable, and including the heteroskedastic uncertainties on both
the dependent and independent variables.
There are two excellent reasons for modelling the relation stochastically;
firstly, if the intrinsic scatter is ignored and the relation between
variables is assumed to be deterministic, those data points with smaller
measurement uncertainties may have an unrepresentative greater weighting
during the fitting process.
Secondly, we are interested in producing probability distributions over stellar
densities and surface gravities, as opposed to point estimates, and propagating
these probability distributions through to subsequent analyses.
Several recent studies have required posterior Probability Distribution
Function (PDF) samples, in order to conduct their (hierarchical)
analyses \citep[e.g.][]{Rogers2014, Foreman-Mackey2014, Angus2015}.
Including observational uncertainties on the independent variable within our
analysis is also important, because ordinary least squares regression methods
performed on data with two-dimensional uncertainties will result in a slope
that is biased towards zero
(e.g., Fuller 1987; Akritas \& Bershady 1996; Fox 1997).
or a demonstration of the effects of neglecting intrinsic scatter and
two-dimensional uncertainties, see Kelly (2007).

The two models describing the relationship between flicker, logg and stellar
density can be written as
\begin{equation}
	\log(g) \sim \mathcal{N}(\mu = \alpha + \beta F_8, \sigma^2 = \
	\sigma^2 + \gamma F_8)
\end{equation}
\label{eq:logg}
and
\begin{equation}
	\rho_* \sim \mathcal{N}(\mu = \delta + \epsilon F_8, \sigma^2 = \
	\sigma^2 + \zeta F_8),
\end{equation}
\label{eq:rho}
where $F_8$ is flicker and $\rho_*$ is stellar density.
The free parameters of the two models are $\alpha$, $\beta$, $\sigma_G$,
$\gamma$, $\delta$, $\epsilon$ and $\zeta$.
The two relationships can be described as Gaussian distributions with means
given by the equation of a straight line, and standard deviations given by
observational uncertainties, plus some dispersion which is a function of the
dependent variable.

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
TBD
}
\label{fig:rhostar}
\end{center}
\end{figure}

%%% logg plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
\caption{
TBD
}
\label{fig:logg}
\end{center}
\end{figure}

>>>>

We model the stochastic relationship between $F_8$, \logg\ and \rhostar,
accounting for the fact that there exists some intrinsic scatter in
the dependent variable, and including the heteroskedastic uncertainties on both
the dependent and independent variables.
There are two excellent reasons for modelling the relation stochastically;
firstly, if the intrinsic scatter is ignored and the relation between
variables is assumed to be deterministic, those data points with smaller
measurement uncertainties may have an unrepresentative greater weighting
during the fitting process.
Secondly, we are interested in producing probability distributions over stellar
densities and surface gravities, as opposed to point estimates, and propagating
these probability distributions through to subsequent analyses.
Several recent studies have required posterior Probability Distribution
Function (PDF) samples, in order to conduct their (hierarchical)
analyses \citep[e.g.][]{rogers:2015, foreman-mackey:2014, angus:2015}.

Including observational uncertainties on the independent variable within our
analysis is also important, because ordinary least squares regression methods
performed on data with two-dimensional uncertainties will result in a slope
that is biased towards zero
(e.g., Fuller 1987; Akritas \& Bershady 1996; Fox 1997).
For a demonstration of the effects of neglecting intrinsic scatter and
two-dimensional uncertainties, see Kelly (2007).

The two models describing the relationship between $F_8$, \logg and \rhostar
density can be written as
\begin{equation}
	\log_{10}(\rho_\star) \sim \mathcal{N}(\mu = \alpha + \beta
	\log_{10}(F_8), \sigma = \sigma_{\rho}),
\end{equation}
\label{eq:rho}
and
\begin{equation}
	\log_{10}(g) \sim \mathcal{N}(\mu = \gamma + \delta \log_{10}(F_8),
	\sigma = \sigma_g)
\end{equation}
\label{eq:logg}
The free parameters of the two models are $\alpha$, $\beta$, $\sigma_{\rho}$,
$\gamma$, $\delta$, and $\sigma_g$.
These relationships can be described as Gaussian distributions with means
given by the equation of a straight line, and standard deviations which
describe the intrinsic scatter about the mean.

In order to account for the two-dimensional observational uncertainties in
this data set, we marginalize over the latent, `true' values of $F_8$,
\rhostar and \logg using the importance sampling method of
\citep{hogg:2010}\footnote{Note that while the pseudo-marginal MCMC method of
\citep{andrieu:2009} provides a truly unbiased estimate of the marginalized
likelhood, we expect any bias introduced by the importance sampling method to
be negligable.}.
The observed values of the variables; $F_8$, \rhostar and \logg can be thought
of as a single `draw' from an underlying (assumed Gaussian) probability
distribution, with mean equal to the `true' value\footnote{We use `true' to
mean the value that would have been observed, given infinitely high
signal-to-noise.} and standard deviation equal to the observational
uncertainty.
We compute the likelihood of the data given the model, marginalized over the
true values, using importance sampling.
We sample from the posterior PDFs of the true values of the variables,
conditioned on the observed values, and compute the likelihood of each of those
samples.
% These posterior PDFs could be produced by the astronomers who provide the
% catalog.
% If posterior samples are made available, they can be used in this stage of
% the inference.
% In the event that they are not provided however, they can be generated.
% For example, the $log(g)$ values used here are point estimates of the
% posterior PDFs of $log(g)$, produced by \citet{chaplin:2014} using
% {\it Kepler} light curves and stellar models.
We would use the posterior samples generated in previous model fitting process
if they were available, however generating new samples is an acceptable
approximation provided the posteriors are Gaussian (or log-Normal) and the
priors used by the previous fitters were uninformative.
Such an approximation {\it will not be necessary} for those studies using
\rhostar or \logg calculated using the newly-calibrated relations
presented here, as long as the posterior samples are published.
After producing new posterior PDF samples we then add up the individual
log-likelihoods of each sample to compute the total marginalized likelihood
for a star.
The marginalized likelihood of the whole dataset is then computed as the sum
of the log-likelihoods of each star.
The importance sampling method allows us to approximate the integral over the
latent variables.

Figures \ref{fig:rhostar} and \ref{fig:logg} show the data with the best-fit
model.
The faint colored lines show the model where the parameters have been drawn
from their posterior PDFs.

%%% rhostar plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/rho_vs_flicker.pdf}
\caption{
Stellar density vs. flicker. This figure shows the model, conditioned on the
data. The solid black line shows the median model values, and the pink lines
are random draws from the posterior PDFs of the model parameters; $\alpha$,
$\beta$ and $\sigma_{\rho}$.
}
\label{fig:rhostar}
\end{center}
\end{figure}

%%% logg plot
\begin{figure}
\begin{center}
\includegraphics[width=8.4cm,angle=0,clip=true]{../figs/logg_vs_flicker.pdf}
\caption{
$\log(g)$ vs. flicker. As in \ref{fig:rhostar} this figure shows
the model, conditioned on the data. The solid black line shows the median model
values, and the blue lines are random draws from the posterior PDFs of the
model parameters; $\gamma$, $\delta$ and $\sigma_g$.
}
\label{fig:logg}
\end{center}
\end{figure}

\begin{table}
\caption{Median parameter values with their 16th and 84th percentile
	 uncertainties.}
\begin{tabular}{lcc}
\hline\hline
    Parameter & Median value \\
    \hline
    $\alpha$         &   \\
    $\beta$          &   \\
    $\sigma_{\rho}$  &   \\
    $\gamma$         &   \\
    $\delta$         &   \\
    $\sigma_{g}$     &   \\
    \hline
\end{tabular}
\end{table}

### 2015-06-05 version1.0/introduction.tex
making some corrections to the text

Accurate stellar characterization plays a vital role for many active research
fields within astronomy.
For example, stellar populations, galactic archaeology, the study of binary
stars, asteroseismology and exoplanet studies all rely on inferences of basic
stellar parameters to varying degrees.
Empirically-derived and reliable estimates are of particular value for
increasing our confidence in the end-product results built upon these inputs.

@@

An alternative approach for inferring model parameters is a ``probabilistic''
one.
Unlike the deterministic case, a single observational input is intepreted to be
caused by a range of possible model parameters, described by a probability
distribution.
This statement is true even in the case of a perfect observation of infinite
signal-to-noise, since our underlying model itself comes with uncertainty.
In recent years there has been a shift towards such methods in several areas of
 astronomy and particularly within the exoplanet community.
For example, \citet{wolfgang:2015} considere the mass-radius relationship
of exoplanets to be probabilistic, since a planet of a particular radius could
have a range of masses thanks to variances in compositions, environment and
other circumstances.
These recent demonstrations in exoplanetary science prompt us to consider the
need for treating the parent stars in the same probabilistic framework, with
potential applications spanning many fields of astronomy.

The demand for probilistic stellar parameters is not only motivated by the fact
that probability distributions are far more representative of our `beliefs'
about astrophysical parameters, it also has a practical purpose.
When using data published in the astronomical literature to, for example, infer
relationships between observed parameters, that inference can be performed as
the final stage in a hierarchical treatment.
This is useful if one wants to account for multi-dimensional uncertainties by
marginalizing over the `true' parameter values upon which the noisy
observations are dependent.
This point is discussed further in \textsection\ref{sec:HBM}.

>>>>

Accurate stellar characterization plays a vital role for many active research
fields within astronomy.
For example, stellar populations, galactic archaeology, the study of binary
stars, asteroseismology and exoplanet studies all rely on inferences of basic
stellar parameters to varying degrees.
% Empirically-derived and reliable estimates are of particular value for
% increasing our confidence in the end-product results built upon these inputs.

@@

An alternative approach for inferring model parameters is a ``probabilistic''
one.
Unlike the deterministic case, a single observational input is intepreted to be
caused by a range of possible model parameters, described by a probability
distribution.
This statement is true even in the case of a perfect observation of infinite
signal-to-noise, since our underlying model itself comes with uncertainty.
In recent years there has been a shift towards such methods in several areas of
 astronomy and particularly within the exoplanet community.
For example, \citet{wolfgang:2015} considered the mass-radius relationship
of exoplanets to be probabilistic, since a planet of a particular radius could
have a range of masses thanks to variances in compositions, environment and
other circumstances.
These recent demonstrations in exoplanetary science prompt us to consider the
need for treating the parent stars in the same probabilistic framework, with
potential applications spanning many fields of astronomy.

The demand for probabilistic stellar parameters is not only motivated by the
fact that probability distributions are far more representative of our
`beliefs' about astrophysical parameters, it also has a practical purpose.
When using data published in the astronomical literature to, for example, infer
relationships between observed parameters, that inference can be performed as
the final stage in a hierarchical treatment.
This is useful if one wants to account for multi-dimensional uncertainties by
marginalizing over the `true' parameter values upon which the noisy
observations are dependent.
This point is discussed further in \textsection\ref{sec:HBM}.

### 2015-06-05 version1.0/manuscript.tex
making some corrections to the text

TBD

@@

One of the more recent tools developed to characterize stars is known as
``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an
8-hour timescale (dubbed $F_8$) in a broad visible bandpass time series
photometric light curve, such as that from \textit{Kepler} or the upcoming TESS
mission. A more detailed account of the proceedure to calculate flicker is
described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker
displays a remarkable correlation to the asteroseismically determined parent
star surface gravities (\logg). Turning this round, the observation implies
that flicker can be used to blindly infer surface gravities at the level of
$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light
curves available through the array of exoplanet transit missions flying and
scheduled to launch.

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
450, 1787
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014}
Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014}
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781,
124
\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
ApJ, 795, 64
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
2166
\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
127
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
Rogers, L.~A., 2015, ApJ, 801, 41
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted
(astro-ph:1504.07557)
\end{thebibliography}

>>>>

% aims
The relations between observable stellar parameters are usually assumed to be
deterministic.
That is, given an infinitely precise measurement of independant
variable, `$x$', and some model, the value of dependant variable, `y' can be
known exactly.
In practise this assumption is rarely valid and intrinsic stochasticity means
that two stars with exactly the same `$x$', will have slightly different
`$y$`s.
% methods
The relation between short-timescale brightness fluctuations (flicker) of stars
and both surface gravity \citep{bastien:2013} and stellar density
\citep{kipping:2014} are two such stochastic relations that have, until now,
been treated as deterministic ones.
We recalibrate these relations in a probabilistic framework, using
Hierarchical Bayesian Modelling (HBM) and constrain the instrinsic scatter in
the relations.
Our treatment also fully accounts for the observational uncertainties on all
variables.
% results
These new relations provide probability distributions over values of $\log(g)$
or \rhostar for given measurements of flicker.
Such probability distributions will be useful for subsequent studies using
probabilistic inference.

@@

One of the more recent tools developed to characterize stars is known as
``flicker'' \citep{bastien:2013}. Flicker is a proxy for the scatter on an
8-hour timescale (dubbed $F_8$) in a broad visible bandpass time series
photometric light curve, such as that from \textit{Kepler} or the upcoming TESS
mission. A more detailed account of the proceedure to calculate flicker is
described in \citet{bastien:2013}. As shown in \citet{bastien:2013}, flicker
displays a remarkable correlation to the asteroseismically determined parent
star surface gravities (\logg). Turning this around, the observation implies
that flicker can be used to blindly infer surface gravities at the level of
$\sim0.1$\,dex, an attractive proposition given the wealth of photometric light
curves available through the array of exoplanet transit missions flying and
scheduled to launch.

@@

% #####################################################################
%% Bibliography
\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Andrieu \& Roberts}{2009}]{andrieu:2009}
Andrieu, C., \& Roberts, G.~O., Ann. Statist., 2, 695
\bibitem[\protect\citeauthoryear{Angus et al.}{2015}]{angus:2015}
Angusr, R., Aigrain, S., Foreman-Mackey, D. \& McQuillan, A., 2015, MNRAS,
450, 1787
\bibitem[\protect\citeauthoryear{Bastien et al.}{2013}]{bastien:2013} Bastien,
F. A., Stassun, K. G., Basri, G. \& Pepper, J., 2013, Nature, 500, 427
\bibitem[\protect\citeauthoryear{Chaplin et al.}{2014}]{chaplin:2014}
Chaplin, W.~J., Basu, S., Huber, D. et al., 2014, ApJS, 210, 1
\bibitem[\protect\citeauthoryear{Cranmer et al.}{2014}]{cranmer:2014}
Cranmer, S. R., Bastien, F. A., Stassun, K. G. \& Saar, S. H., 2014, ApJ, 781,
124
\bibitem[\protect\citeauthoryear{Foreman-Mackey et al.}{2014}]
{foreman-mackey:2014} Foreman-Mackey, D., Hogg, D.~W. \& Morton, T.~D., 2014,
ApJ, 795, 64
\bibitem[\protect\citeauthoryear{Hogg et al.}{2010}]{hogg:2010}
Hogg, D.~W., Myers, A.~D. \& Bovy, J., 2010, ApJ, 725,
2166
\bibitem[\protect\citeauthoryear{Huber et al.}{2013}]{huber:2013}
Huber, D., Chaplin, W.~J., Christensen-Dalsgaard, J. et al., 2013, ApJ, 767,
127
\bibitem[\protect\citeauthoryear{Kipping et al.}{2014}]{kipping:2014}
Kipping, D.~M., Bastien, F.~A., Stassun, K.~G., Chaplin, W.~J., Huber, D. \&
Buchhave, L.~A., 2014, ApJ, 785, 32
\bibitem[\protect\citeauthoryear{Rogers}{2015}]{rogers:2015}
Rogers, L.~A., 2015, ApJ, 801, 41
\bibitem[\protect\citeauthoryear{Wolfgang et al.}{2015}]{wolfgang:2015}
Wolfgang, A., Rogers, L.~A., Ford, E.~B., 2015, ApJ, submitted
(astro-ph:1504.07557)
\end{thebibliography}
